Building socket-aware BPF programs

### 2018-10-27 paper/paper.tex
paper: Minor wording suggestions

\usepackage{xcolor}
\newcommand\todo[1]{\textcolor{red}{#1}}
\begin{filecontents*}{sk-lookup-api.c}
\import{sk-lookup-api.c}
\end{filecontents*}

@@

    To build such awareness into a BPF helper, the verifier needs the ability
    to track the safety of the calls, including appropriate reference counting
    upon the underlying socket. This paper describes extensions to the verifier
    to perform tracking of references in a BPF program. This allows BPF
    developers to extend the UAPI with functions that allocate and release
    resources within the execution lifetime of a BPF program, and the verifier
    will validate that the resources are released exactly once prior to program
    completion.

@@

\begin{itemize}
\item BPF started with sockets, but only provided access to packets
\item Previous helper extensions allowed e.g. socket operation handling
\item Arbitrary socket lookup from socket/packet hooks allows more powerful BPF
      programs
\item Use cases revolve around building logic in socket/packet layer hooks that
      change the behaviour based on the existence and attributes of a socket.
\end{itemize}

@@

\lstinputlisting[float=*t,%
  caption={BPF API helper functions for socket lookup},%
  language=c,%
  label=lst:helper-api,%
  breaklines=true]{sk-lookup-api.c}

>>>>

\usepackage{xcolor}
\newcommand\todo[1]{\textcolor{red}{#1}}

@@

    This paper presents the changes necessary to extend socket introspection
    capabilities of the Linux kernel's BPF packet hooks. To provide safety
    guarantees around the access of the socket, the verifier is extended to
    track allocation and release of kernel resources. Using this mechanism, we
    introduce socket lookup and release helpers into the BPF API.

    %These new primitives open up new possibilities for various use cases
    %including stateful firewalling and load balancing.

    %To provide access to socket state from a BPF helper call, the verifier
    %needs the ability to track the safety of the calls, including appropriate
    %reference counting upon the underlying socket for the duration of the
    %access.

    %With such an ability ability in the verifier, we add socket lookup and
    %release function calls to the BPF API, allowing BPF programs to safely find
    %a socket and build logic upon the presence or attributes of a socket.
    %Applications for this functionality range from load-balancing traffic based
    %on the presence of a listening application to building stateful firewalling
    %primitives to understand whether traffic for this connection has been seen
    %before. With this new functionality, BPF programs can integrate more
    %closely with the networking stack's understanding of the traffic transiting
    %the kernel.

    %provide safe access to sockets from BPF packet handling hooks. The BPF verifier
    %is extended to gain understanding of reference-counted pointer types, which
    %allows BPF code paths to

@@

\begin{itemize}
\item BPF started with sockets, but only provided access to packets
\item Previous helper extensions allowed e.g. socket operation handling
\item Arbitrary socket lookup from socket/packet lookups allows more powerful
      BPF programs
\end{itemize}

% Over the past several years, BPF has steadily become more powerful in
% multiple ways: Through building more intelligence into the verifier which
% allows more complex programs to be loaded, and through extension of the API
% such as by adding new map types and new native BPF function calls. While
% BPF has its roots in applying filters at the socket layer, the ability to
% introspect the sockets relating to traffic being filtered has been limited.

To build such awareness into a BPF helper, the verifier needs the ability to
track the safety of the calls, including appropriate access into socket pointer
types, and reference counting upon the underlying socket. This paper walks
through extensions to the verifier to perform tracking of references in a BPF
program. This allows BPF developers to extend the API with functions that
allocate and release resources within the execution lifetime of a BPF program,
and the verifier will validate that the resource acquisition and release is
balanced across all execution paths of the BPF programs.

\todo{Desire to target logic based upon socket existence/attributes at packet
layers}

@@

The consideration of each of these items reflects the definition of the lookup
helpers (Listing~\ref{lst:helper-api}):

\todo{Span this listing across two columns}
\begin{lstlisting}[caption={BPF API helper functions for socket lookup},language=c,label=lst:helper-api,breaklines=true]
struct bpf_sock *bpf_sk_lookup_tcp(void *ctx, struct bpf_sock_tuple *tuple, u32 tuple_size,
                                   u32 netns, u64 flags)
struct bpf_sock *bpf_sk_lookup_udp(void *ctx, struct bpf_sock_tuple *tuple, u32 tuple_size,
                                   u32 netns, u64 flags)
int bpf_sk_release(struct bpf_sock *sk)
\end{lstlisting}
